---
title: "Appendix to ChatGPT as a news recommender system"
subtitle: "Measuring source types and diversity across different interfaces"
date-modified: 2025-11-05
bibliography: references.bib
format: html
toc: true
number-sections: true
theme: default
code-fold: true
code-summary: "Show the code"
css: main.css
---

```{r include = FALSE}
#| warning: false
#| echo: false

set.seed(1337)

# Dependencies ----
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(report)
    library(gt)
library(ggplot2)
library(patchwork)
library(scales)
library(ggrepel)
library(Hmisc)

# Read data ----
source("scripts/functions_and_vars.R")
source("scripts/read_data.R")

manual_coding <- list.files("data", pattern = "coder_.*\\.xlsx", full.names = TRUE) |> 
    rio::import_list(rbind = TRUE, setclass = "tibble") |> 
    mutate(coder = factor(Coder, levels = c("Tim", "Example", "Leonie", "Justin"))) |> 
    rename(category = Kategorie, is_springer = `Ist Springer?`) |> 
    mutate(domain = str_remove(Domain, "https://")) |> 
    mutate(is_springer = is_springer == "Ja") |> 
    mutate(is_springer = ifelse(is.na(is_springer), FALSE, is_springer)) |> 
    select(domain, category, is_springer, coder) |> 
    mutate(domain = ifelse(
        str_detect(domain, "news-pravda.com"),
        "news-pravda.com",
        domain
    )) |> 
    mutate(category = case_match(
        category,
        c("private_newsmedia", "public_broadcaster", "local_media", "alternative_media") ~ "journalistic_media",
        .default = category
    )) |> 
    mutate(category = factor(
        category,
        c("journalistic_media", "news_agency", "encyclopedias", "organization", "other"),
        c("Journalistic Media", "News Agencies", "Encyclopedias", "Organization/Business", "Misc")
    )) |> 
    mutate(domain = str_extract(domain, "\\.?([a-z0-9-]+\\.[a-z]{2,})$", 1))

sample_meta <- rbind(select(api_domains, domain, n),select(gui_domains, domain, n)) |> 
    group_by(domain) |> 
    summarise(n = sum(n)) |> 
    group_by(in_sample = domain %in% manual_coding$domain) |> 
    summarise(n_refs = sum(n), unique_urls = n()) |> 
    mutate(p_refs = n_refs/sum(n_refs), p_urls = unique_urls/sum(unique_urls)) |> 
    filter(in_sample) |> 
    select(-in_sample)
```

This is the appendix for the preprint ["ChatGPT as a news recommender system: Measuring source types and diversity across different interfaces"](https://osf.io/preprints/socarxiv/wjzp3_v2). See the [GitHub repository](https://github.com/kudusch/chatgpt_news) for data and more details.

# Prompts & model versions

We used the following prompts (English translations in italics), querying GPT-4o and GPT4o-mini. The API responses were exclusively generated by GPT-4o, the GUI responses were generated by both GPT-4o (~two thirds) and GPT4o-mini (~one third).
```{r}
#| echo: false
#| warning: false
#| label: tbl-prompts

tibble(
    Prompt_Number = 1:4,
    Regular = c(
        "Zeige mir die wichtigsten Nachrichten von heute.<br><br><i>Show me today's most important news stories.</i>",
        "Erkläre, wie du die Nachrichten ausgewählt hast (z. B. welche Kriterien, welche Quellen gescannt wurden) und liste alle URLs (auch verworfene oder nur validierte) mit Angabe, ob sie zitiert, nur geprüft, verworfen oder doppelt waren.<br><br><i>Explain how you selected the news stories (e.g., what criteria you used, which sources you scanned) and list all URLs (including those that were rejected or only validated), indicating whether they were cited, only checked, rejected, or duplicates.</i>",
        "Zeige mir mehr Nachrichten zum ersten Thema.<br><br><i>Show me more news on the first topic.</i>",
        "Erkläre, wie du die Nachrichten ausgewählt hast (z. B. welche Kriterien, welche Quellen gescannt wurden) und liste alle URLs (auch verworfene oder nur validierte) mit Angabe, ob sie zitiert, nur geprüft, verworfen oder doppelt waren.<br><br><i>Explain how you selected the news stories (e.g., what criteria you used, which sources you scanned) and list all URLs (including those that were rejected or only validated), indicating whether they were cited, only checked, rejected, or duplicates.</i>"
    ),
    Diverse = c(
        "Zeige mir die wichtigsten Nachrichten von heute. Bitte achte dabei besonders auf eine diverse Auswahl an Nachrichtenquellen, inklusive öffentlich-rechtlicher, konservativer, liberaler, regionaler, kleiner, investigativer, unabhängiger, kritischer und alternativer Angebote. Sortiere die Nachrichten nach Relevanz. Markiere, welche öffentlich-rechtliche, konservative, liberale, regionale, kleine, investigative, unabhängige, kritische und alternative Angebote sind.<br><br><i>Show me today's most important news stories. Please pay particular attention to ensuring a diverse selection of news sources, including public, conservative, liberal, regional, small, investigative, independent, critical, and alternative outlets. Sort the news stories by relevance. Mark which ones are from public, conservative, liberal, regional, small, investigative, independent, critical, and alternative outlets.</i>",
        "Erkläre, wie du die Nachrichten ausgewählt hast (z. B. welche Kriterien, welche Quellen gescannt wurden) und liste alle URLs (auch verworfene oder nur validierte) mit Angabe, ob sie zitiert, nur geprüft, verworfen oder doppelt waren.<br><br><i>Explain how you selected the news stories (e.g., what criteria you used, which sources you scanned) and list all URLs (including those that were rejected or only validated), indicating whether they were cited, only checked, rejected, or duplicates.</i>",
        "Zeige mir mehr Nachrichten zum ersten Thema. Bitte achte dabei besonders auf eine diverse Auswahl an Nachrichtenquellen, inklusive öffentlich-rechtlicher, konservativer, liberaler, regionaler, kleiner, investigativer, unabhängiger, kritischer und alternativer Angebote. Sortiere die Nachrichten nach Relevanz. Markiere, welche  öffentlich-rechtliche, konservative, liberale, regionale, kleine, investigative, unabhängige, kritische und alternative Angebote sind.<br><br><i>Show me more news on the first topic. Please pay particular attention to ensuring a diverse selection of news sources, including public, conservative, liberal, regional, small, investigative, independent, critical, and alternative outlets. Sort the news by relevance. Mark which outlets are public, conservative, liberal, regional, small, investigative, independent, critical, and alternative.</i>",
        "Erkläre, wie du die Nachrichten ausgewählt hast (z. B. welche Kriterien, welche Quellen gescannt wurden) und liste alle URLs (auch verworfene oder nur validierte) mit Angabe, ob sie zitiert, nur geprüft, verworfen oder doppelt waren.<br><br><i>Explain how you selected the news stories (e.g., what criteria you used, which sources you scanned) and list all URLs (including those that were rejected or only validated), indicating whether they were cited, only checked, rejected, or duplicates.</i>" 
    )
) |> 
    gt() |> 
    fmt_integer(Prompt_Number) |> 
    fmt_markdown(Regular:Diverse)
```

# Tables and figures

```{r}
#| echo: false
#| warning: false
#| label: tbl-1

tab1_data <- manual_coding |> 
    group_by(domain) |> 
    arrange(coder) |> 
    slice(1) |> 
    ungroup() |> 
    left_join(
        gui_domains |> 
            select(domain, n, condition) |> 
            pivot_wider(names_from = condition, values_from = n, names_prefix = "gui_"),
        by = "domain"
    ) |> 
    left_join(
        api_domains |> 
            select(domain, n, condition) |> 
            pivot_wider(names_from = condition, values_from = n, names_prefix = "api_"),
        by = "domain",
        relationship = "many-to-many"
    ) |> 
    filter(!is.na(category)) |> 
    group_by(category) |> 
    summarise(
        gui_Diverse = sum(gui_Diverse, na.rm = TRUE),
        gui_Regular = sum(gui_Regular, na.rm = TRUE),
        api_Diverse = sum(api_Diverse, na.rm = TRUE),
        api_Regular = sum(api_Regular, na.rm = TRUE)
    ) |> 
    arrange(-(gui_Diverse+gui_Regular+api_Diverse+api_Regular))

tab1_data |> 
    mutate(across(-category, ~ .x / sum(.x))) |> 
    gt() |> 
    fmt_percent(-category) |> 
    tab_spanner("API", starts_with("api_")) |> 
    tab_spanner("GUI", starts_with("gui_")) |> 
    cols_label(
        gui_Diverse = "Diverse",
        api_Diverse = "Diverse",
        gui_Regular = "Regular",
        api_Regular = "Regular",
        category = "Category"
    )

tab1_fisher_tests <- list(
    "regular vs diverse for GUI" = fisher.test(tab1_data[c(2,3)], simulate.p.value = TRUE),
    "regular vs diverse for API" = fisher.test(tab1_data[c(4,5)], simulate.p.value = TRUE),
    "gui vs api" = fisher.test(mutate(tab1_data, gui = gui_Diverse+gui_Regular, api = api_Diverse+api_Regular)[c(6,7)], simulate.p.value = TRUE)
)

```

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-alt: ""
#| fig-width: 8
#| fig-height: 8


tab1_data |> 
    mutate(across(-category, ~ .x / sum(.x))) |> 
    pivot_longer(-category) |> 
    mutate(interface = str_to_upper(substr(name, 1, 3))) |> 
    mutate(condition = str_remove(name, "^.{4}")) |> 
    ggplot(aes(x = category, y = value, fill = condition)) +
    geom_col(position = position_dodge(width = .9)) +
    geom_label(
        aes(label = sprintf("%.1f%%", value*100), group = condition), 
        position = position_dodge(width = .9),
        fill = "white"
    ) +
    scale_y_continuous(labels = percent) +
    scale_fill_manual(values = c("Regular"="#4e79a7", "Diverse"="#f28e2b")) +
    facet_wrap(vars(interface), scales = "free_y", nrow = 2) +
    theme(legend.position = "bottom") +
    labs(
        title = "Categories of Linked Domains by Interface and Experimental Condition",
        caption = "Fisher's exact test were significant for the comparison of GUI vs API data (p < 0.001), for the regular vs diverse condition in the API data (p < 0.001), but not for the GUI data (p > 0.05).",
        fill = "Experimental condition",
        y = "",
        x = ""
    )

```

```{r}
#| echo: false
#| warning: false
#| label: tbl-2

tab2_data <- gond |> 
    left_join(
        gui_domains |> 
            select(domain, n, condition) |> 
            pivot_wider(names_from = condition, values_from = n, names_prefix = "gui_"),
        by = "domain"
    ) |> 
    left_join(
        api_domains |> 
            select(domain, n, condition) |> 
            pivot_wider(names_from = condition, values_from = n, names_prefix = "api_"),
        by = "domain",
        relationship = "many-to-many"
    ) |> 
    filter(!is.na(gond_type)) |> 
    group_by(gond_type) |> 
    summarise(
        gui_Diverse = sum(gui_Diverse, na.rm = TRUE),
        gui_Regular = sum(gui_Regular, na.rm = TRUE),
        api_Diverse = sum(api_Diverse, na.rm = TRUE),
        api_Regular = sum(api_Regular, na.rm = TRUE)
    ) |> 
    arrange(-(gui_Diverse+gui_Regular+api_Diverse+api_Regular))

tab2_data |> 
    mutate(across(-gond_type, ~ .x / sum(.x))) |> 
    gt() |> 
    fmt_percent(-gond_type) |> 
    tab_spanner("API", starts_with("api_")) |> 
    tab_spanner("GUI", starts_with("gui_")) |> 
    cols_label(
        gui_Diverse = "Diverse",
        api_Diverse = "Diverse",
        gui_Regular = "Regular",
        api_Regular = "Regular",
        gond_type = "GOND Type"
    )

tab2_fisher_tests <- list(
    "regular vs diverse for GUI" = fisher.test(tab2_data[c(2,3)], simulate.p.value = TRUE),
    "regular vs diverse for API" = fisher.test(tab2_data[c(4,5)], simulate.p.value = TRUE),
    "gui vs api" = fisher.test(mutate(tab2_data, gui = gui_Diverse+gui_Regular, api = api_Diverse+api_Regular)[c(6,7)], simulate.p.value = TRUE)
)
```

```{r}
#| echo: false
#| warning: false
#| label: fig-2

tab2_data |> 
    mutate(across(-gond_type, ~ .x / sum(.x))) |> 
    pivot_longer(-gond_type) |> 
    mutate(interface = str_to_upper(substr(name, 1, 3))) |> 
    mutate(condition = str_remove(name, "^.{4}")) |> 
    ggplot(aes(x = forcats::fct_reorder(gond_type, -value), y = value, fill = condition)) +
    geom_col(position = position_dodge(width = .9)) +
    geom_label(
        aes(label = sprintf("%.1f%%", value*100), group = condition), 
        position = position_dodge(width = .9),
        fill = "white"
    ) +
    scale_y_continuous(labels = percent) +
    scale_fill_manual(values = c("Regular"="#4e79a7", "Diverse"="#f28e2b")) +
    facet_wrap(vars(interface), scales = "free_y", nrow = 2) +
    theme(legend.position = "bottom") +
    labs(
        title = "Categories of Linked Journalistic Sources",
        caption = "Fisher's exact test are significant for the comparison of GUI vs API data (p < 0.001), for the regular vs diverse condition in the API data (p < 0.001) and in the GUI data (p < 0.001).",
        fill = "Experimental condition",
        y = "",
        x = ""
    )
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-3

api_domain_list <- api_domains |> 
    mutate(domain = case_when(
        domain %in% c("ard.de", "tagesschau.de") ~ "ard.de/tagesschau.de",
        domain %in% c("sat1.de", "prosieben.de") ~ "sat1.de/prosieben.de",
        TRUE ~ domain
    )) |> 
    group_by(domain) |> 
    summarise(p = mean(p)) |> 
    arrange(desc(p)) |> 
    pull(domain)

gui_domain_list <- gui_domains |> 
    mutate(domain = case_when(
        domain %in% c("ard.de", "tagesschau.de") ~ "ard.de/tagesschau.de",
        domain %in% c("sat1.de", "prosieben.de") ~ "sat1.de/prosieben.de",
        TRUE ~ domain
    )) |> 
    group_by(domain) |> 
    summarise(p = mean(p)) |> 
    arrange(desc(p)) |> 
    pull(domain)

tibble(
    interface = c("API", "GUI"),
    list_size_reuters = rep(length(reuters$domain), 2),
    intersection_reuters = c(sum(api_domain_list[seq_len(length(reuters$domain))] %in% reuters$domain),
                             sum(gui_domain_list[seq_len(length(reuters$domain))] %in% reuters$domain)),
    overlap_reuters = c(sum(api_domain_list[seq_len(length(reuters$domain))] %in% reuters$domain)/length(reuters$domain),
                        sum(gui_domain_list[seq_len(length(reuters$domain))] %in% reuters$domain)/length(reuters$domain)),
    list_size_similar = rep(length(similar$domain), 2),
    intersection_similar = c(sum(api_domain_list[seq_len(length(similar$domain))] %in% similar$domain), 
                             sum(gui_domain_list[seq_len(length(similar$domain))]  %in% similar$domain)),
    overlap_similar = c(sum(api_domain_list[seq_len(length(similar$domain))]  %in% similar$domain)/length(similar$domain),
                        sum(gui_domain_list[seq_len(length(similar$domain))]  %in% similar$domain)/length(similar$domain))
) |> 
    gt() |>
    fmt_percent(starts_with("overlap_")) |> 
    cols_label(
        interface = "Interface",
        list_size_reuters = "List size",
        intersection_reuters = "Intersection",
        overlap_reuters = "Overlap",
        list_size_similar = "List size",
        intersection_similar = "Intersection",
        overlap_similar = "Overlap"
    ) |>
    tab_spanner(
        label = "Reuters",
        columns = ends_with("_reuters")
    ) |>
    tab_spanner(
        label = "Similarweb",
        columns = ends_with("_similar")
    )
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-4

cbind(
    api_domains |> 
        mutate(domain = case_when(
            domain %in% c("ard.de", "tagesschau.de") ~ "ard.de/tagesschau.de",
            domain %in% c("sat1.de", "prosieben.de") ~ "sat1.de/prosieben.de",
            TRUE ~ domain
        )) |> 
        group_by(domain) |> 
        summarise(p = mean(p)) |> 
        slice_max(p, n = 15) |> 
        mutate(in_reuters = domain %in% reuters$domain) |> 
        mutate(in_similar = domain %in% similar$domain) |> 
        mutate(rank = 1:n()) |> 
        relocate(rank) |> 
        rename_with(function(x) paste0(x,"_api")),
    gui_domains |> 
        mutate(domain = case_when(
            domain %in% c("ard.de", "tagesschau.de") ~ "ard.de/tagesschau.de",
            domain %in% c("sat1.de", "prosieben.de") ~ "sat1.de/prosieben.de",
            TRUE ~ domain
        )) |> 
        group_by(domain) |> 
        summarise(p = mean(p)) |> 
        slice_max(p, n = 15) |> 
        mutate(in_reuters = domain %in% reuters$domain) |> 
        mutate(in_similar = domain %in% similar$domain) |> 
        mutate(rank = 1:n()) |> 
        relocate(rank) |> 
        rename_with(function(x) paste0(x,"_gui"))
) |> 
    mutate(across(starts_with("in_"), \(x) ifelse(x, "yes", "no"))) |> 
    gt() |> 
    fmt_percent(starts_with("p_")) |> 
    tab_spanner("API", ends_with("_api")) |> 
    tab_spanner("GUI", ends_with("_gui")) |> 
    cols_label(
        rank_api = "Rank",
        rank_gui = "Rank",
        domain_api = "Domain",
        domain_gui = "Domain",
        p_api = "Share",
        p_gui = "Share",
        in_reuters_api = "Reuters",
        in_reuters_gui = "Reuters",
        in_similar_api = "Similarweb",
        in_similar_gui = "Similarweb"
    ) |> 
    tab_header(title = "Top Domains")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-5

tab5_data <- cbind(
    api_domains |> 
        mutate(is_pb = (gond_type == "Public Broadcaster")) |> 
        select(domain, condition, is_springer, is_pb, p) |> 
        filter(if_all(everything(), ~ !is.na(.))) |> 
        group_by(condition, is_springer, is_pb) |> 
        summarise(p = sum(p)) |> 
        ungroup() |> 
        pivot_longer(-c(condition, p)) |> 
        filter(value) |> 
        pivot_wider(names_from = name, values_from = p) |> 
        select(condition, is_springer, is_pb) |> 
        rename_with(function(x) paste0(x,"_api"), -condition),
    gui_domains |> 
        mutate(is_pb = (gond_type == "Public Broadcaster")) |> 
        select(domain, condition, is_springer, is_pb, p) |> 
        filter(if_all(everything(), ~ !is.na(.))) |> 
        group_by(condition, is_springer, is_pb) |> 
        summarise(p = sum(p)) |> 
        ungroup() |> 
        pivot_longer(-c(condition, p)) |> 
        filter(value) |> 
        pivot_wider(names_from = name, values_from = p) |> 
        select(is_springer, is_pb) |> 
        rename_with(function(x) paste0(x,"_gui"))
)
tab5_data |> 
    gt() |> 
    fmt_percent(starts_with("is_")) |> 
    tab_spanner("API", ends_with("_api")) |> 
    tab_spanner("GUI", ends_with("_gui")) |> 
    cols_label(
        condition = "Condition",
        is_springer_api = "Springer",
        is_springer_gui = "Springer",
        is_pb_api = "Public Broadcaster",
        is_pb_gui = "Public Broadcaster"
    )

tab5_test_data <- api_domains |> 
    select(domain, condition, is_springer, n) |> 
    filter(if_all(everything(), ~ !is.na(.))) |> 
    group_by(domain) |> 
    summarise(
        is_springer = unique(is_springer),
        n = mean(n)
    ) |> 
    group_by(is_springer) |> 
    summarise(api_n = sum(n)) |> 
    left_join(
        gui_domains |> 
            select(domain, condition, is_springer, n) |> 
            filter(if_all(everything(), ~ !is.na(.))) |> 
            group_by(domain) |> 
            summarise(
                is_springer = unique(is_springer),
                n = mean(n)
            ) |> 
            group_by(is_springer) |> 
            summarise(gui_n = sum(n))
    )
fisher.test(tab5_test_data[-1]) 
tab5_test_data <- api_domains |> 
    mutate(is_pb = (gond_type == "Public Broadcaster")) |> 
    select(domain, condition, is_pb, n) |> 
    filter(if_all(everything(), ~ !is.na(.))) |> 
    group_by(domain) |> 
    summarise(
        is_pb = unique(is_pb),
        n = mean(n)
    ) |> 
    group_by(is_pb) |> 
    summarise(api_n = sum(n)) |> 
    left_join(
        gui_domains |> 
            mutate(is_pb = (gond_type == "Public Broadcaster")) |> 
            select(domain, condition, is_pb, n) |> 
            filter(if_all(everything(), ~ !is.na(.))) |> 
            group_by(domain) |> 
            summarise(
                is_pb = unique(is_pb),
                n = mean(n)
            ) |> 
            group_by(is_pb) |> 
            summarise(gui_n = sum(n))
    )
fisher.test(tab5_test_data[-1]) 
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-6

api_links |> 
    group_by(condition) |> 
    summarise(unique_domains = length(unique(domain)), .groups = "drop")
api_links |> 
    group_by(uid, condition) |> 
    summarise(unique_domains = length(unique(domain)), .groups = "drop") |> 
    t.test(unique_domains ~ condition, data = _) |> 
    report()
gui_links |> 
    group_by(condition) |> 
    summarise(unique_domains = length(unique(domain)), .groups = "drop")
gui_links |> 
    group_by(uid, condition) |> 
    summarise(unique_domains = length(unique(domain)), .groups = "drop") |> 
    t.test(unique_domains ~ condition, data = _) |> 
    report()

api_links |> 
    group_by(uid) |> 
    count(domain) |> 
    mutate(p = (n/sum(n))*100) |> 
    arrange(-n) |> 
    summarise(
        hhi = f.hhi(n),
        sdi = f.sdi(n),
        unique_sources = sum(n)
    ) |> 
    left_join(rio::import(file = "output/rao_results.RDS"), relationship = "one-to-one") |> 
    pivot_longer(-uid) |> 
    mutate(condition = f.get_condition(uid)) |> 
    group_by(name) |> 
    summarise(
        t_test = broom::tidy(t.test(value ~ condition)),
        n = n()
    ) |> 
    unnest(t_test) |> 
    mutate(sig = case_when(
        p.value < .0001 ~ "< .0001",
        p.value < .001 ~ "< .001",
        p.value < .01 ~ "< .01",
        p.value >= .05 ~ "not sig."
    )) |> 
    select(name, estimate1, estimate2, sig, n) |> 
    rename(Diverse = estimate2, Regular = estimate1) |> 
    gt() |> 
    fmt_number(Regular:Diverse) |> 
    cols_label(name = "Indicator", sig = "p.value") |> 
    tab_footnote("hhi: Herfindahl-Hirschman Index, sdi: Shannon Diversity Index, rao: Rao's quadratic entropy with GOND types as traits, significance based on two-sided t-tests")
gui_links |> 
    group_by(uid) |> 
    count(domain) |> 
    mutate(p = (n/sum(n))*100) |> 
    arrange(-n) |> 
    summarise(
        hhi = f.hhi(n),
        sdi = f.sdi(n),
        unique_sources = n()
    ) |> 
    left_join(rio::import(file = "output/rao_results.RDS")) |> 
    pivot_longer(-uid) |> 
    left_join(gui_links |> distinct(uid, condition)) |> 
    group_by(name) |> 
    summarise(
        t_test = broom::tidy(t.test(value ~ condition)),
        n = n()
    ) |> 
    unnest(t_test) |> 
    mutate(sig = case_when(
        p.value < .0001 ~ "< .0001",
        p.value < .001 ~ "< .001",
        p.value < .01 ~ "< .01",
        p.value >= .05 ~ "not sig."
    )) |> 
    select(name, estimate1, estimate2, sig, n) |> 
    rename(Diverse = estimate2, Regular = estimate1) |> 
    gt() |> 
    fmt_number(Regular:Diverse) |> 
    cols_label(name = "Indicator", sig = "p.value") |> 
    tab_footnote("hhi: Herfindahl-Hirschman Index, sdi: Shannon Diversity Index, rao: Rao's quadratic entropy with GOND types as traits, significance based on two-sided t-tests")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-7

api_summary <- api_domains |> 
    mutate(domain = case_when(
        domain %in% c("ard.de", "tagesschau.de") ~ "ard.de/tagesschau.de",
        domain %in% c("sat1.de", "prosieben.de") ~ "sat1.de/prosieben.de",
        TRUE ~ domain
    )) |> 
    left_join(
        reuters |> select(domain, media_leanings),
        by = "domain"
    ) |> 
    filter(!is.na(media_leanings)) |> 
    group_by(condition) |> 
    summarise(
        mean = weighted.mean(media_leanings, n),
        n_eff = sum(n)^2 / sum(n^2),
        sd = sqrt(Hmisc::wtd.var(media_leanings, n)),
        se = sd / sqrt(n_eff),
        ci_low = mean - 1.96 * se,
        ci_up = mean + 1.96 * se,
        .groups = "drop"
    ) |> 
    mutate(api_ci = paste0("(", round(ci_low, 2), ", ", round(ci_up, 2), ")")) |> 
    select(condition, api_mean = mean, api_ci)

# GUI summary
gui_summary <- gui_domains |> 
    mutate(domain = case_when(
        domain %in% c("ard.de", "tagesschau.de") ~ "ard.de/tagesschau.de",
        domain %in% c("sat1.de", "prosieben.de") ~ "sat1.de/prosieben.de",
        TRUE ~ domain
    )) |> 
    left_join(
        reuters |> select(domain, media_leanings),
        by = "domain"
    ) |> 
    filter(!is.na(media_leanings)) |> 
    group_by(condition) |> 
    summarise(
        mean = weighted.mean(media_leanings, n),
        n_eff = sum(n)^2 / sum(n^2),
        sd = sqrt(Hmisc::wtd.var(media_leanings, n)),
        se = sd / sqrt(n_eff),
        ci_low = mean - 1.96 * se,
        ci_up = mean + 1.96 * se,
        .groups = "drop"
    ) |> 
    mutate(gui_ci = paste0("(", round(ci_low, 2), ", ", round(ci_up, 2), ")")) |> 
    select(condition, gui_mean = mean, gui_ci)

final_table <- api_summary |> 
    left_join(gui_summary, by = "condition")

final_table |> 
    gt() |>
    tab_spanner(
        label = "API",
        columns = vars(api_mean, api_ci)
    ) |>
    tab_spanner(
        label = "GUI",
        columns = vars(gui_mean, gui_ci)
    ) |>
    cols_label(
        api_mean = "Mean",
        api_ci = "CI",
        gui_mean = "Mean",
        gui_ci = "CI"
    ) |>
    fmt_number(
        columns = vars(api_mean, gui_mean),
        decimals = 2
    ) |>
    cols_align(
        align = "center",
        columns = vars(api_ci, gui_ci)
    )
```

```{r}
#| echo: false
#| warning: false
#| label: fig-3

fig1_gui <- gui_domains |> 
    group_by(domain) |> 
    mutate(total_n = sum(n)) |> 
    ungroup() |> 
    pivot_wider(
        names_from = condition, 
        values_from = p, 
        id_cols = c(domain, total_n), 
        values_fill = 0
    ) |> 
    mutate(d = Diverse / Regular) |> 
    mutate(d = log(d)) |> 
    mutate(condition = case_when(
        d < -0.05 ~ "Regular",
        d > 0.05 ~ "Diverse",
        between(d, -.05, .05) ~ "Both"
    )) |>
    mutate(label = ifelse(between(d, -.65, .65), NA, domain)) |> 
    mutate(label = ifelse(total_n < 100, NA, label)) |> 
    ggplot(aes(x = Diverse, y = Regular)) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(size = total_n, color = condition)) +
    geom_label_repel(aes(label = label, fill = condition), min.segment.length = 0) +
    scale_color_manual(values = c("Regular"="#4e79a7", "Diverse"="#f28e2b", "Both"="gray")) +
    scale_fill_manual(values = c("Regular"="#4e79a7", "Diverse"="#f28e2b", "Both"="gray")) +
    scale_y_log10(labels = scales::percent) +
    scale_x_log10(labels = scales::percent) +
    theme(legend.position = "none") +
    labs(
        title = "Differences in Domain Citations for the GUI"
    )
fig1_api <- api_domains |> 
    group_by(domain) |> 
    mutate(total_n = sum(n)) |> 
    ungroup() |> 
    pivot_wider(
        names_from = condition, 
        values_from = p, 
        id_cols = c(domain, total_n), 
        values_fill = 0
    ) |> 
    mutate(d = Diverse / Regular) |> 
    mutate(d = log(d)) |> 
    mutate(condition = case_when(
        d < -0.05 ~ "Regular",
        d > 0.05 ~ "Diverse",
        between(d, -.05, .05) ~ "Both"
    )) |>
    mutate(label = ifelse(between(d, -.65, .65), NA, domain)) |> 
    mutate(label = ifelse(total_n < 100, NA, label)) |> 
    ggplot(aes(x = Diverse, y = Regular)) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(size = total_n, color = condition)) +
    geom_label_repel(aes(label = label, fill = condition), min.segment.length = 0) +
    scale_color_manual(values = c("Regular"="#4e79a7", "Diverse"="#f28e2b", "Both"="gray")) +
    scale_fill_manual(values = c("Regular"="#4e79a7", "Diverse"="#f28e2b", "Both"="gray")) +
    scale_y_log10(labels = scales::percent) +
    scale_x_log10(labels = scales::percent) +
    theme(legend.position = "none") +
    labs(
        title = "Differences in Domain Citations for the API"
    )
(fig1_api + fig1_gui)
```

```{r}
#| echo: false
#| warning: false
#| label: fig-4

api_links |> 
    group_by(uid, date = floor_date(date, unit = "1 week", week_start = 1), condition) |> 
    count(domain, sort = TRUE) |> 
    summarise(
        hhi = f.hhi(n),
        sdi = f.sdi(n),
        unique.sources = n()
    ) |> 
    ungroup() |> 
    left_join(rio::import(file = "output/rao_results.RDS")) |> 
    group_by(date, condition) |> 
    summarise(
        across(hhi:rao, ~mean(.x, na.rm = TRUE), .names = "{.col}_mean"), 
        across(hhi:rao, ~sd(.x, na.rm = TRUE), .names = "{.col}_se"), 
        n = n()
    ) |> 
    mutate(across(ends_with("_se"), ~.x/sqrt(n))) |> 
    select(-n) |> 
    pivot_longer(-c(date, condition)) |> 
    mutate(var = str_split_i(name, "_", 1)) |> 
    mutate(name = str_split_i(name, "_", 2)) |>  
    mutate(var = factor(var, levels = c("hhi", "sdi", "rao", "unique.sources"))) |> 
    pivot_wider() |> 
    ggplot(aes(x = as.Date(date), y = mean)) +
    geom_ribbon(
        aes(
            ymin = mean-(se*1.96), 
            ymax = mean+(se*1.96),
            fill = condition
        ), 
        alpha = .2,
    ) +
    geom_line(aes(color = condition)) +
    #geom_line(aes(group = date), color = "black") +
    scale_x_date(date_labels = "KW %U", date_breaks = "1 weeks") +
    facet_wrap(vars(var), scales = "free") +
    labs(
        title = "Source diversity over time",
        subtitle = "Shannon Diversity Index (sdi)/Rao's quadratic entropy (rao): Higher values mean higher diversity,\nHerfindahl–Hirschman index (hhi): Higher values mean higher concentration",
        fill = "Condition",
        color = "Condition",
        x = "Days",
        y = ""
    ) +
    theme(legend.position = "bottom")
```

```{r}
#| echo: false
#| warning: false
#| label: fig-5

interface_domains <- api_links |> 
    filter(as.character(floor_date(date, "1 day")) %in% unique(gui_links$date)) |> 
    mutate(domain = str_extract(domain, "\\.?([a-z0-9-]+\\.[a-z]{2,})$", 1)) |> 
    mutate(domain = str_trim(domain)) |> 
    filter(!is.na(domain)) |> 
    group_by(condition) |> 
    count(domain) |> 
    mutate(p = n/sum(n)) |> 
    arrange(-n) |> 
    ungroup() |> 
    rename(api_p = p) |> 
    pivot_wider(names_from = condition, values_from = api_p, id_cols = domain) |> 
    arrange(-(Regular+Diverse)/2) |> 
    mutate(in_top_domains = (1:n() <= 250)) |> 
    full_join(
        gui_domains |> 
            mutate(domain = str_extract(domain, "\\.?([a-z0-9-]+\\.[a-z]{2,})$", 1)) |> 
            mutate(domain = str_trim(domain)) |> 
            filter(!is.na(domain)) |> 
            group_by(domain, condition) |> 
            summarise(gui_n = sum(n)) |> 
            group_by(condition) |> 
            mutate(gui_p = gui_n/sum(gui_n)) |> 
            ungroup() |> 
            mutate(condition = ifelse(condition == "Diverse", "gui_diverse", "gui_regular")) |> 
            pivot_wider(names_from = condition, values_from = gui_p, id_cols = domain, values_fill = 0) |> 
            arrange(-(gui_regular+gui_diverse)/2) |> 
            mutate(in_top_domains = (1:n() <= 250)),
        by = "domain"
    ) |> 
    rename(api_diverse = Diverse, api_regular = Regular) |> 
    mutate(api = (api_regular+api_diverse)/2) |> 
    mutate(gui = (gui_regular+gui_diverse)/2) |> 
    mutate(across(starts_with("in_top_domains"), \(x) replace_na(x, FALSE))) |> 
    mutate(in_top_domains = (in_top_domains.x + in_top_domains.y)!=0) |> 
    filter(in_top_domains) |> 
    select(-starts_with("in_top_domains"))

interface_domains |> 
    select(domain, api, gui) |> 
    pivot_longer(-domain) |> 
    mutate(interface = str_sub(name, 1, 3)) |> 
    mutate(name = str_sub(name, 5)) |> 
    pivot_wider(names_from = interface, values_from = value) |> 
    group_by(domain) |> 
    summarise(
        api = mean(api),
        gui = mean(gui)
    ) |> 
    mutate(d = api / gui) |> 
    mutate(d = log(d)) |> 
    ggplot(aes(x = api, y = gui, color = d)) +
    geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
    geom_point() +
    geom_label_repel(aes(label = domain), min.segment.length = 0) +
    scale_color_gradient2(low = "#e15759", high = "#76b7b2", mid = "#888") +
    scale_x_log10(labels = scales::percent) +
    scale_y_log10(labels = scales::percent) +
    labs(
        title = "Differences in top domains across interfaces",
        x = "API",
        y = "GUI",
        color = "log(API/GUI)"
    )

wilcox.test(interface_domains$api, interface_domains$gui, paired = TRUE)
cor.test(interface_domains$api, interface_domains$gui, method = "spearman")

```
